---
title: "What YouTube videos make the trending section in the United States?"
output: html_document
---

```{r setup, echo = FALSE, message = F, include = F}
library("knitr")
library("dplyr")
source("scripts/doughnut-graph.R")
source("scripts/daily_views_graph.R")
```
## Introduction

Trending YouTube videos capture what is going on in the world through a variety
of different categories. YouTube videos are something that almost everyone in 
the United States is familiar with. However, the types of videos and **the 
logistics behind the videos that make the trending section are not as 
well known**. In this report we will analyze what factors support videos making 
it into YouTube's trending section _through summaries, tables, and visual 
representations_ of our findings from our data set. The data set we will be
using to conduct this analysis was created and organized by 
[Rishav Sharma.](https://www.kaggle.com/rsrishav/youtube-trending-video-dataset?select=US_youtube_trending_data.csv)
We pulled the data set on November 14th, 2020 and thus the data we are analyzing
contains values from August 12th, 2020 to November 14th, 2020. It is important
to note that our analysis is only for videos **in the United States** and does 
not contain correlations between the factors we are analyzing and other 
countries.

## Summary Information


## Charts

### Doughnut Graph
This doughnut graph displays the distribution of each video category found 
within the trending section. From looking at the graph, it is clear to see that
the majority of videos that make it into the trending section are **Music, 
Entertainment, and Sports**, with the combination of the three making up over
50% of the fifteen unique categories that were found in our data.

```{r, echo = FALSE }
trending_categories_plot
```

### Average Viewership Per Day

```{r, echo = FALSE }
days_of_week_viewership_plot
```