---
title: "What YouTube videos make the trending section in the United States?"
output: html_document
---

```{r setup, echo = FALSE, message = F, include = F}
### Reading in file
youtube_trending <- read.csv("data/US_youtube_trending_data.csv",
  encoding = "UTF-8",
  stringsAsFactors = FALSE
)
library("knitr")
library("dplyr")
source("scripts/piechart-graph.R")
source("scripts/daily_views_graph.R")
source("scripts/boxplot-graph.R")
source("scripts/summary_info_script.R")
source("scripts/Grouped_by.R")
```
## Introduction

Trending YouTube videos capture what is going on in the world through a variety
of different categories. YouTube videos are something that almost everyone in 
the United States is familiar with. However, the types of videos and **the 
logistics behind the videos that make the trending section are not as 
well known**. In this report we will analyze what factors support videos making 
it into YouTube's trending section _through summaries, tables, and visual 
representations_ of our findings from our data set. The data set we will be
using to conduct this analysis was created and organized by 
[Rishav Sharma.](https://www.kaggle.com/rsrishav/youtube-trending-video-dataset?select=US_youtube_trending_data.csv)
We pulled the data set on November 14th, 2020 and thus the data we are analyzing
contains values from August 3rd, 2020 to November 14th, 2020. It is important
to note that our analysis is only for videos **in the United States** and does 
not contain correlations between the factors we are analyzing and other 
countries.

## Summary Information
```{r include = FALSE}
summary_info <- get_summary_info(youtube_trending)
```
Our dataset has `r summary_info$rows` rows and `r summary_info$columns` columns. 

The earliest date in our dataset is `r summary_info$earliest_date` and 
the latest recorded date is `r summary_info$latest_date`.

These are the column names: `r summary_info$column_names`.

## Aggregate Table
For the table, we grouped by categoryId from the dataset to see which category 
trends the most, and other details that might come with it. We found out that 
**Gaming** came out as the most trending category with the highest average 
views. As for the title stat, we added that to see if having a video title in
all caps can influence the ability to trends as shown on the table.
```{r, fig.align='center',out.extra='angle=90', echo = FALSE, message = FALSE}
kable(get_table_info(youtube_trending),
      col.names = c("CATEGORY", "AVERAGE VIEWS", "AVERAGE LIKES", "TOP CHANNEL",
                    "AVERAGE COMMENTS", "TITLE STATISTICS"),
      align = 'c')
```

## Charts

### Trending Categories by Percentage
This pie chart displays the distribution of each video category found within the
trending section as a percentage. Each section of the pie represents a 
proportion to the total number of videos within the trending section during the 
3 month period we pulled our data from. When hovering over a specific section of
the graph you are provided with the category name, the number of videos that 
fall under that specific category from our data, and the percentage of the whole
rounded to two decimal places that represents how frequent a particular
category makes it into the trending section. We created this chart to get an 
idea of what videos make it onto YouTube's trending section more frequently than
others.

```{r, fig.align='center',out.extra='angle=90', echo = FALSE, message = FALSE, warning = FALSE}
trending_categories_graph(youtube_trending)
```

From looking at the graph, it is clear to see that the majority of videos that 
make it into the trending section are **Music, Entertainment, and Sports**, with
the combination of the three making up over 50% of the fifteen unique categories
that were found in our data. An interesting bit of information that I found was
that YouTube has more than just 15 unique categories, but the trending section
seems to group several categories into a main category to be used in their
trending section. 

### Average Viewership Per Day
This plot displays the average amount of views trending videos recieved on a given day (Sunday to Saturday). We created this chart to see if there is a relationship between average viewership and the day of the week. 

```{r, fig.align='center',out.extra='angle=90', echo=FALSE, message=FALSE}
get_daily_views_plot(youtube_trending)
```

This graph clearly shows that trending videos on Friday receive nearly triple the average viewership. Additionally, Wednesday is a particularly slow day for YouTube videos. This may mean that popular creators should attempt to have a video on the trending page every Friday because this is when they would likely receive the highest viewership.

### Amount of Time for a Video to Trend
This graph shows the variation of the time it takes a video to reach the
trending page after it’s published. The graph provides us with the mean, median,
min, and max values of how long it took a video to trend and is grouped by 
category.

Most of the data is concentrated between the 0 to 5 day range, while there are 
some outliers above 10 days for most categories. Interestingly enough, the data
also included values with negative values. This is because the published date
includes the full date and time, whereas the trending date only includes the 
month and day. For videos to have a negative value of days it took to reach the 
trending page, it means that the video was wildly popular enough to be trending 
on the same date it was uploaded. When plotting the graph, I felt that it was 
best to replace the negative time values with a 0 in order to keep consistency 
because negative time values are not technically correct.
```{r, fig.align='center',out.extra='angle=90', echo = FALSE, message = FALSE, warning = FALSE}
time_until_trending_graph(youtube_trending)
```
Based on the information provided by the graph, the Music category seems to have
an overwhelming amount of outliers that began trending far after the video was 
published. This is most likely due to the fact that new music takes time to 
circulate around before it becomes popular on mainstream media.