---
title: "What YouTube videos make the trending section in the United States?"
output: html_document
---

```{r setup, echo = FALSE, message = F, include = F}
### Reading in file
youtube_trending <- read.csv("data/US_youtube_trending_data.csv",
  encoding = "UTF-8",
  stringsAsFactors = FALSE
)
library("knitr")
library("dplyr")
source("scripts/piechart-graph.R")
source("scripts/daily_views_graph.R")
source("scripts/boxplot-graph.R")
```
## Introduction

Trending YouTube videos capture what is going on in the world through a variety
of different categories. YouTube videos are something that almost everyone in 
the United States is familiar with. However, the types of videos and **the 
logistics behind the videos that make the trending section are not as 
well known**. In this report we will analyze what factors support videos making 
it into YouTube's trending section _through summaries, tables, and visual 
representations_ of our findings from our data set. The data set we will be
using to conduct this analysis was created and organized by 
[Rishav Sharma.](https://www.kaggle.com/rsrishav/youtube-trending-video-dataset?select=US_youtube_trending_data.csv)
We pulled the data set on November 14th, 2020 and thus the data we are analyzing
contains values from August 12th, 2020 to November 14th, 2020. It is important
to note that our analysis is only for videos **in the United States** and does 
not contain correlations between the factors we are analyzing and other 
countries.

## Summary Information


## Charts

### Trending Categories by Percentage
This pie chart displays the distribution of each video category found 
within the trending section. From looking at the graph, it is clear to see that
the majority of videos that make it into the trending section are **Music, 
Entertainment, and Sports**, with the combination of the three making up over
50% of the fifteen unique categories that were found in our data.

```{r, fig.align='center',out.extra='angle=90', echo = FALSE, warning = F}
trending_categories_plotly
```

### Average Viewership Per Day
This plot displays the average amount of views trending videos recieved on a given day (Sunday to Saturday). We created this chart to see if there is a relationship between average viewership and the day of the week. 

```{r, fig.align='center',out.extra='angle=90', echo=FALSE, message=FALSE}
get_daily_views_plot(youtube_trending)
```

This graph clearly shows that trending videos on Friday receive nearly triple the average viewership. Additionally, Wednesday is a particularly slow day for YouTube videos. This may mean that popular creators should attempt to have a video on the trending page every Friday because this is when they would likely receive the highest viewership.

### Boxplot Graph
This graph shows the variation of the time it takes a video to reach the
trending page after it's published. Most of the data is concentrated between the
0 to 5 day range, while there are some outliers above 10 days for most 
categories. Interestingly enough, the graph also included values with negative 
values. This is because the published date includes the full date and time, 
whereas the trending date only includes the month and day. For videos to have a 
negative value of days it took to reach the trending page, it means that the 
video was wildly popular enough to be trending on the same date it was uploaded.

```{r, fig.align='center',out.extra='angle=90', echo = FALSE }
days_until_trending_plotly
```